---
title: "FGFR2-centered PPI network plot and modulatory analysis"
author: "Kelin G. deOliveira, Mauro A. A. Castro"
date: "`r Sys.Date()`"
output:  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{FGFR2-centered PPI network plot and modulatory analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\SweaveUTF8
---
# Plotting FGFR2 interactome

##Load necessary tools

###Packages
```{r eval=FALSE}
library(RedeR)

library(igraph)

library(RColorBrewer)
```
### Data-mining results
```{r eval=FALSE}
data("Body_PPI")

data("Breast_PPI")
```

##Call Java interface for RedeR
```{r eval=FALSE}
rdp<- RedPort()

calld(rdp)
```

## Tissue-inespecific connections
Create variables containing the vertices and edges. Convert data frame to an igraph object.
```{r eval=FALSE}
vert_fgfr2_ints<- Body_PPI[,c(2,17)]
edge_fgfr2_ints<- Body_PPI[,c(1,2)]
colnames(edge_fgfr2_ints)<- c("from", "to")

gr_fgfr2_ints<-graph_from_data_frame(edge_fgfr2_ints, directed = T, vertices = vert_fgfr2_ints)
```

Choose nodeColor pallete to receive values from the "Total" column.
```{r eval=FALSE}
rcb<- brewer.pal(9, "BuPu")
color_col <- colorRampPalette(rcb)(13)

gr_fgfr2_ints<- att.setv(g=gr_fgfr2_ints, from="Total", to="nodeColor", cols = color_col)
```

Plot the igraph object into the RedeR's Java interface, add legend and organize plot with relax function.
```{r eval=FALSE}
addGraph(rdp, gr_fgfr2_ints, gcoord=c(50,50), gscale=110, theme= 'tm1')
addLegend.color(rdp, colvec=gr_fgfr2_ints, title="Databases")
relax(rdp,p1=10,p2=80,p3=10,p4=100,p5=100,p6=100,p7=10,p8=10,ps=FALSE)
```

Select nodes that represent connections that occured 10 times or more.
```{r eval=FALSE}
nodes<- as.character(Body_PPI$Int.2[which(Body_PPI$Total>=10)])
selectNodes(rdp,nodes)
```


## Breast tissue-specific connections
Create variables containing the vertices and edges. Convert data frame to an igraph object.
```{r eval=FALSE}
vert_ppi_m<- Breast_PPI[,c(2,8)]
edge_ppi_m<- Breast_PPI[,c(1,2)]
colnames(edge_ppi_m)<- c("from", "to")
gr_ppi_m<-graph_from_data_frame(edge_ppi_m, directed = T, vertices = vert_ppi_m)
```

Choose nodeColor pallete to receive values from the "Total" column.
```{r eval=FALSE}
rcb<- brewer.pal(9, "BuPu")
color_col <- colorRampPalette(rcb)(13)

gr_ppi_m<- att.setv(g=gr_ppi_m, from="Total", to="nodeColor", cols = color_col)
```

Plot the igraph object into the RedeR's Java interface, add legend and organize plot with relax function.
```{r eval=FALSE}
addGraph(rdp, gr_ppi_m, gcoord=c(50,50), gscale=110, theme= 'tm1')
addLegend.color(rdp, colvec=gr_ppi_m, title="Databases")
relax(rdp,p1=10,p2=80,p3=10,p4=100,p5=100,p6=100,p7=10,p8=10,ps=FALSE)
```

Select nodes that represent connections that occured 3 times or more.
```{r eval=FALSE}
nodes<- as.character(Breast_PPI$Int.2[which(Breast_PPI$Total>=3)])
selectNodes(rdp,nodes)
```

Clean environment
```{r eval=FALSE}
rm(list = ls())
```


# Modulatory analysis (Mindy algorithm)

## Load necessary tools
### Packages
```{r eval=FALSE}
library(RTN)
library(Fletcher2013b)
```

### Datasets from Fletcher, M. et al., 2013
```{r eval=FALSE}
data("rtni1st")
data("rtni2nd")
rtni1st<- upgradeTNI(rtni1st)
rtni2nd<- upgradeTNI(rtni2nd)
```

### Data-mining results
```{r eval=FALSE}
data("Breast_PPI")
```

```{r eval=FALSE}
# Obtain all gene symbols in rtni1st or rtni2n
prots_1st<- as.vector(rtni1st@rowAnnotation$SYMBOL)
prots_2nd<- as.vector(rtni2nd@rowAnnotation$SYMBOL)

# Selection, removal of duplicated and concatenation of gene symbols to be used as candidate
# modulators (those that interact with fgfr2)
data4mod<- c(as.character(Breast_PPI$Int.2))
data4mod<- data4mod[!duplicated(data4mod)]
modulators_1st<- prots_1st[prots_1st%in%data4mod]
modulators_2nd<- prots_2nd[prots_2nd%in%data4mod]

# Retrieve PROBEIDs for each modulator
idx_1st<- which(prots_1st%in%data4mod)
mod_probeid_1st<- rtni1st@annotation$PROBEID[idx_1st]
idx_2nd<- which(prots_2nd%in%data4mod)
mod_probeid_2nd<- rtni2nd@annotation$PROBEID[idx_2nd]
```

## Selection of Transcription Factors
```{r eval=FALSE}
# TF clusters related with breast cancer risk (CASTRO, M. et al., 2015)
TFs<- c("SNAPC2", "MZF1", "RARA", "AR", "SPDEF", "ZNF587", "ARNT2", "BRD8", "MYB",
        "ESR1", "GATA3", "YPEL3", "FOXA1",  "AFF3", "XBP1", "ZNF552", "NFIB", "YBX1",
        "TBX19", "SOX10", "CEBPB", "TRIM29", "CBFB")
```
## Conditional Analysis
```{r eval=FALSE}
# With cohort one
cond_1st<- tni.conditional(rtni1st, modulators = mod_probeid_1st, tfs = TFs,
                           pValueCutoff = 1e-3)
res_1st<- tni.get(cond_1st, what = "cdtrev")

# With cohort two
cond_2nd<- tni.conditional(rtni2nd, modulators = mod_probeid_2nd, tfs = TFs,
                           pValueCutoff = 1e-3)
res_2nd<- tni.get(cond_2nd, what = "cdtrev")
```
## Results
```{r eval=FALSE}
mods_1st<-list()
mods_1st$PROBEID<-names(unlist(lapply(res_1st, nrow)))
idx_mods<- which(rtni1st@rowAnnotation$PROBEID %in% mods_1st$PROBEID)
mods_1st$SYMBOL<-rtni1st@rowAnnotation$SYMBOL[idx_mods]

mods_2nd<-list()
mods_2nd$PROBEID<-names(unlist(lapply(res_2nd, nrow)))
idx_mods<- which(rtni2nd@rowAnnotation$PROBEID %in% mods_2nd$PROBEID)
mods_2nd$SYMBOL<-rtni2nd@rowAnnotation$SYMBOL[idx_mods]
```

```{r eval=FALSE}
# save in a RData
save(cond_1st, cond_2nd, file = "./FGFR2_Conditional_TNIs.RData")
save(res_1st, res_2nd, file = "./FGFR2_Conditional.RData")
save(mods_1st, mods_2nd, file= "./FGFR2_Conditional_Annotation.RData")
```
